Think deep for reef resilience

### 2025-08-16 paper/paper_draft.qmd
Add equation cross reference to get label
Align to "predictors" over "features" and "characteristics"

$$ O:I = \frac{(S_O/N_O)}{(S_I/N_I)} $$

@@

Partial Dependence Analysis was conducted to understand the relationship between individual
features and time series cluster predictions using a trained Random Forest (RF) classifier
[@friedman2001]. The analysis utilised the reef characteristics of interest including the
median reef depth, mean DHW for the given GCM, total connectivity strength, and the
outgoing-to-incoming connectivity ratio (\$O:I\$). The full dataset consisted of the
timeseries cluster assignment of each of the 2429 reefs under the five GCMs explored, for a
total of 12,145 samples. The dataset was randomly split into training (60%) and testing
(40%) subsets to ensure model generalisability. The RF classifier achieved a 99% accuracy on
the training set and 66% on the test set.

@@

Partial dependence analysis was performed by applying the trained model to the test dataset.
Values for individual reef characteristic were systematically (one-at-a-time) varied across
a grid of 50 equally spaced points spanning its range, while all other features were left at
their modelled values. The procedure creates (1) a modified test dataset, where the target
reef characteristic was set to the grid value for all samples, (2) generates probabilistic
predictions for all three cluster classes using the trained RF classifier, and (3) the mean
predicted probability for each class was calculated across all test observations. The median
probability was also assessed (not shown) but was found not to affect the interpretation of
the analysis results. The process isolates the marginal effect of each feature by averaging
over the natural distribution of all other features in the test dataset.

>>>>

$$
O:I = \frac{(S_O/N_O)}{(S_I/N_I)}
$$ {#eq-so-to-si}

@@

Partial Dependence Analysis was conducted to understand the relationship between individual
features and time series cluster predictions using a trained Random Forest (RF) classifier
[@friedman2001]. The analysis utilised the reef characteristics of interest as the
predictors, including the median reef depth, mean DHW for the given GCM, total connectivity
strength, and the outgoing-to-incoming connectivity ratio (\$O:I\$). The full dataset
consisted of the timeseries cluster assignment of each of the 2429 reefs under the five GCMs
explored, for a total of 12,145 samples. The dataset was randomly split into training (60%)
and testing (40%) subsets to ensure model generalisability. The RF classifier achieved a 99%
accuracy on the training set and 66% on the test set.

@@

Partial dependence analysis was performed by applying the trained model to the test dataset.
Values for individual predictors were systematically (one-at-a-time) varied across a grid of
50 equally spaced points spanning its range, while all other features were left at their
modelled values. The procedure creates (1) a modified test dataset, where the target
predictor was set to the grid value for all samples, (2) generates probabilistic predictions
for all three cluster classes using the trained RF classifier, and (3) the mean predicted
probability for each class was calculated across all test observations. The median was also
assessed (not shown) but was found not to substantially influence the interpretation of the
analysis results. The process isolates the marginal effect of each predictor by averaging
over the natural distribution of all other representative reef characteristics in the test
dataset.
